Manuscript

### 2022-02-03 analysis/manuscript.Rmd
wflow_publish("analysis/manuscript.Rmd")

options(scipen=999)
source("code/settings.R")
source("R/functions.R")
library(targets)
library(kableExtra)
library(dplyr)
library(DT)
library(knitr)
library(ggplot2)
library(kableExtra)
library(cowplot)
library(plotly)
library(plyr)
```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(couple_MR_vs_trait_corr)
tar_load(corr_impact_by_traits)
tar_load(traits_corr2_filled)
tar_load(household_MR_summary_AM)
tar_load(num_tests_by_PCs)
tar_load(num_tests_by_PCs_AM_sig)
tar_load(household_MR_binned_het_sex_spec)
tar_load(household_MR_binned_het_joint)
tar_load(traits_final)
tar_load(household_MR_AM_FvsM_fig)
tar_load(confounder_traits)

@@

# Analyses update.

## MR effects vs. raw correlation.

Sought to compare the raw correlation amongst couples vs. the standardized MR effects. There are many traits where the `correlation > MR effects` including some traits of interest such as standing height, place of birth, among many others. A plot comparing the two and a summary table of the effects is shown below.

@@

plot_data <- replace_Neale_ID(couple_MR_vs_trait_corr, traits_corr2_filled, "exposure_ID", "outcome_description") %>% 
  dplyr::rename(Phenotype = outcome_description) %>% 
  dplyr::rename(Correlation = couple_r) %>% dplyr::rename(MR_est = IVW_beta_std)

plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = Correlation,
                                                       y = MR_est, Phenotype = Phenotype)) +
  geom_point(alpha = 3/4) +

  geom_abline(slope=1, intercept=0) +

  ggplot2::geom_errorbar(
    ggplot2::aes(ymin = MR_est - IVW_se_std, ymax = MR_est + IVW_se_std),
    colour = "grey",
    width = 0
  ) +

  theme_minimal() +
  ggplot2::labs(x = paste0("Couple correlation"),
                y = paste("Standardized MR effect"))

ggplotly(plot, tooltip = c("x", "y", "Phenotype"))

@@

@@

plot_data.DT <- plot_data %>% dplyr::rename(IVW_beta_std = MR_est) %>% 
  mutate(mr_corr_diff = Correlation - IVW_beta_std) %>% 
  mutate(correlation_larger = ifelse(Correlation > IVW_beta_std, TRUE, FALSE)) %>%
  dplyr::select(Phenotype, Correlation, IVW_beta_std, IVW_pval_std, mr_corr_diff, correlation_larger)

@@

## Same-trait analyses.

**A few key changes to the pipeline since last time we looked at these results:**

1. List of traits of interest changed to exclude dietary traits and add binary traits (among a few other minor revisions), total number of traits went down. 
2. Sample changed to ensure both males and females were not related (previously had only removed unrelated within males and females separately), our resulting samples closely matched Neale round 2 GWAS which resulted in much fewer couples (~50K instead of ~65K).
3. Filtered SNPs for evidence of reverse causality (should not have any bearing on these results since these involve only one trait).
4. Meta-analyses are performed at SNP-level rather than MR-level.

### Re-calculate number of significant results in same trait analyses.

@@

```
- The number of analyses corresponds to the number of traits (`r dim(traits_final)[1]`). 
- The significant results after adjusting for multiple hypothesis testing are shown below. 
- `IVW_meta_beta` and `IVW_meta_pval` corresponds to the beta and p-value of the meta-analyzed MR across sexes, respectively (i.e. MR estimates were computed in each sex-seperately using sex-specific SNP-exposure and SNP-outcome results and then meta-analyzed).  
- After adjusting for multiple tests (p < `0.05/`r num_tests_by_PCs``), identified `r dim(AM_result_sig)[1]` significant assortative mating MR results (corresponding to the table below).

@@

### Sex heterogeneity.

@@

Figures compared the male to female MR results, and vice versa are shown below. *Suggests evidence of regression dilution bias?*

```{r sex_het_fig, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
household_MR_AM_FvsM_fig[[1]]
household_MR_AM_FvsM_fig[[2]]
```

### Binned results.

@@

slope_sig_age <- household_MR_binned_het_joint %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "age_even_bins") %>% 
  filter(bin_slope_pval < 0.05/num_tests_by_PCs_AM_sig)

slope_sig_tt <- household_MR_binned_het_joint %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "time_together_even_bins") %>% 
  filter(bin_slope_pval < 0.05/num_tests_by_PCs_AM_sig)

slope_wt_sig_age <- household_MR_binned_het_joint %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "age_even_bins") %>% 
  filter(bin_slope_pval_wt < 0.05/num_tests_by_PCs_AM_sig)

slope_wt_sig_tt <- household_MR_binned_het_joint %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "time_together_even_bins") %>% 
  filter(bin_slope_pval_wt < 0.05/num_tests_by_PCs_AM_sig)

@@

```{r binned_results, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

full_binn <- rbind.fill(household_MR_binned_het_sex_spec, household_MR_binned_het_joint) %>% filter(same_trait)
full_binn.DT <- full_binn %>% dplyr::select(exposure_ID, exposure_sex, grouping_var, IVW_beta, IVW_pval, sex_het_pval, bin_slope_beta,
                                            bin_slope_pval,bin_slope_sex_het_pval, bin_slope_sex_het_pval_wt) %>% unique() %>% arrange(exposure_ID)

full_binn.DT <- replace_Neale_ID(full_binn.DT, traits_corr2_filled, "exposure_ID", "outcome_description")

format_round_cols <- colnames(dplyr::select_if(full_binn.DT, is.numeric))

DT::datatable(full_binn.DT, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)

@@

### Summary.

- From the same-trait analyses we unforutantely are not left with many significant results after adjusting for multiple tests. 
- The plot comparing male to female estimates does seem to be an effect of regression dilution bias, since the inverse plot suggests the inverse conclusion? Am I interpretting this right? 
- Our most significant results are the binned results by age, where we have 4 traits that show significant differences according to the slope (the result is the same regardless of using the weighted or unweighted beta estimates), but none of the joint results are BF significant. The four traits are: `r sex_het_wt_sig_age_traits`.

## Confounder analysis with added variables.

Repeated the confounder analysis with the following 4 traits:

```{r confounder_traits, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
confounder_trait_tibble <- as_tibble(confounder_traits)
confounder_trait_tibble_descript <- replace_Neale_ID(confounder_trait_tibble, traits_corr2_filled, "value", "outcome_description")
confounder_trait_tibble_descript %>% 
  kbl() %>%
  kable_styling()
```

Below are the corresponding figures:

@@

@@

  plot_data <- corr_impact_by_traits %>% filter(trait_interest==trait) %>% filter(trait_interest!=trait_j)
  plot_data <- replace_Neale_ID(plot_data, traits_corr2_filled, "trait_j", "Phenotype")
  plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = trait_j_couple_corr,
                                                         y = corr_due_to_confounding, Phenotype = Phenotype)) +
    geom_point(alpha = 3/4) +

    geom_smooth(mapping = aes(x = trait_j_couple_corr, y = corr_due_to_confounding), method = "lm", se=FALSE, formula = y~x+0, fullrange=TRUE, color = custom_col[4]) +
    theme_minimal() +
    ggplot2::labs(x = paste0("Couple trait correlation"),
                  y = paste0("Couple correlation due to confounding"),
                  title = paste0("Impact of ", trait_description, " on couple correlation"))

@@

ggplotly(plots[[1]], tooltip = c("x", "y", "Phenotype"))
ggplotly(plots[[2]], tooltip = c("x", "y", "Phenotype"))
ggplotly(plots[[3]], tooltip = c("x", "y", "Phenotype"))
ggplotly(plots[[4]], tooltip = c("x", "y", "Phenotype"))

@@

>>>>

options(scipen=999)
source("code/settings.R")
source("R/functions.R")
library(targets)
library(kableExtra)
library(DT)
library(knitr)
library(ggplot2)
library(kableExtra)
library(cowplot)
library(plotly)
library(plyr)
library(dplyr)
library(ggrepel)
library(viridis)
library(hrbrthemes)
library(ggpubr)
library(broom)
```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(couple_MR_vs_trait_corr)

tar_load(corr_impact_by_PCs)
tar_load(corr_impact_by_traits)
tar_load(corr_impact_by_coords)
tar_load(corr_potential_trait_confounders)
tar_load(traits_corr)
tar_load(PCs_corr)
tar_load(traits_corr2_filled)
tar_load(household_MR_summary_AM)
tar_load(num_tests_by_PCs)
tar_load(num_tests_by_PCs_AM_sig)
tar_load(household_MR_binned_het_sex_spec)
tar_load(household_MR_binned_het_SNPmeta)
tar_load(traits_final)
tar_load(household_MR_AM_FvsM_fig)
tar_load(confounder_traits)
tar_load(proxyMR_prod_comparison_fig_yiyp_adj)
tar_load(proxyMR_comparison_summary_yiyp_adj_SNPmeta)
tar_load(household_MR_summary_AM)
tar_load(household_MR_binned_het_SNPmeta)

@@

# A few notes (for myself).

- The number of effective tests can be found at the target `num_tests_by_PCs`. 
- The number of effective tests among those significant in the MR can be found at the target `num_tests_by_PCs_AM_sig`.

# Relationship between causal effects and raw phenotypic correlation in couples.

Sought to compare the raw correlation amongst couples vs. the standardized MR effects. There are many traits where the `correlation > MR effects` including some traits of interest such as standing height, place of birth, among many others. A plot comparing the two and a summary table of the significant effects is shown below.

@@

interesting_phenotypes <- c("Basal metabolic rate", "Place of birth in UK - north co-ordinate", 
                            "Systolic blood pressure, automated reading", "Standing height", 
                            "Body fat percentage", "Forced vital capacity (FVC), Best measure","Red blood cell (erythrocyte) count")

plot_data <- replace_Neale_ID(couple_MR_vs_trait_corr, traits_corr2_filled, "exposure_ID", "outcome_description") %>% 
  mutate(Phenotype = outcome_description) %>% 
  mutate(Phenotype = case_when(Phenotype == "Basal metabolic rate" ~ "BMR",
                               Phenotype == "Systolic blood pressure, automated reading" ~ "SBP",
                               Phenotype == "Body fat percentage" ~ "BFP",
                               Phenotype == "Forced vital capacity (FVC), Best measure" ~ "FVC",
                               Phenotype == "Red blood cell (erythrocyte) count" ~ "RBC count",
                               Phenotype == "Place of birth in UK - north co-ordinate" ~ "Place of birth, NC",
                               TRUE ~ Phenotype)) %>%
  dplyr::rename(Correlation = couple_r) %>% dplyr::rename(MR_est = IVW_beta) %>% 
  mutate(diff_p_sig = case_when(diff_p < 0.05/num_tests_by_PCs ~ TRUE, TRUE ~ FALSE)) %>%
  mutate(label_phenos = case_when(outcome_description %in% interesting_phenotypes ~ TRUE, TRUE ~ FALSE))

  # TRUE - BF sig difference 
  # FALSE - non-sig 
  #mutate(diff_p_sig = ifelse(diff_p < 0.05/num_tests_by_PCs, TRUE, FALSE))

plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = Correlation,
                                                       y = MR_est, 
                                                       label = Phenotype, 
                                                       color = factor(diff_p_sig)),) +
  geom_point(alpha = 1.2) +
  scale_color_manual(values = custom_col[c(1,2)], labels=c("Non-significant","BF-signifncant difference")) +
  #geom_text(show.legend = FALSE) +
  geom_abline(slope=1, intercept=0) +

  ggplot2::geom_errorbar(
    ggplot2::aes(ymin = MR_est - IVW_se, ymax = MR_est + IVW_se),
    colour = "grey",
    width = 0
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(xmin = Correlation - couple_r_se, xmax = Correlation + couple_r_se),
    colour = "grey",
    width = 0
  ) +

  theme_minimal() +
  ggplot2::labs(x = paste0("Couple correlation"),
                y = paste("Standardized MR effect")) +
  theme(legend.position="bottom") + theme(legend.title=element_blank()) +
  geom_text_repel(data = subset(plot_data, label_phenos==TRUE),
                  nudge_y = -0.25 - subset(plot_data, label_phenos==TRUE)$MR_est, size = 2,
                  direction = "x")

plot

plot_data.DT <- plot_data %>% dplyr::rename(IVW_beta = MR_est) %>% 
  filter(diff_p < 0.05/num_tests_by_PCs) %>%
  mutate(mr_corr_diff = Correlation - IVW_beta) %>% 
  mutate(correlation_larger = ifelse(Correlation > IVW_beta, TRUE, FALSE)) %>%
  dplyr::select(outcome_description, Correlation, IVW_beta, IVW_pval, diff_p, mr_corr_diff)

@@

## Summary of cases where correlation > MR estimate.

The table below displays the `r dim(plot_data.DT)[1]` traits that have correlation significantly > MR-estimate.

@@

@@

## Search for potential confounders to explain cases where correlation > MR estimate.

1. Restrict to only traits that have correlation significantly > MR-estimate (`r dim(plot_data.DT)[1]` traits have correlation > MR_estimate).
2. For each of these traits, find the traits which have a causal effect on this trait (MR p-val < 0.05/[number of effective tests]).
3. Calculate the correlation due to confounding as: $C = (\alpha_{z\rightarrow x})^2 * \alpha_{z_i\rightarrow z_p}$.
4. Calculate the ratio of this correlation ($C$) and the correlation of $X$ in partners ($C/r_x$).

A summary of the these results can be seen below:

```{r confounder_summary, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

number_of_potential_confounders <- dim(corr_potential_trait_confounders)[1]/dim(plot_data.DT)[1]

non_sig_confounders <- corr_potential_trait_confounders %>% filter(!sig_confounder) %>% group_by(outcome_ID) %>%
  dplyr::summarize(description = outcome_description[1], n = n()) %>% 
  filter(n==number_of_potential_confounders) %>%
  mutate(n = n-number_of_potential_confounders) %>% mutate(potential_confounders = NA) %>%
  mutate(mean_C_ratio = NA) %>%
  dplyr::rename(trait_ID = outcome_ID)

sig_confounders <- corr_potential_trait_confounders %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
  #mutate(potential_confounders = paste0(exposure_description, collapse = "; ")) %>%
  dplyr::summarize(description = outcome_description[1], n = n(), 
                   potential_confounders = paste0(exposure_description, collapse = "; "), 
                   mean_C_ratio = mean(corr_due_to_confounding_ratio)) %>%
  dplyr::rename(trait_ID = outcome_ID)

confounder_summary <- rbind(non_sig_confounders, sig_confounders)

format_round_cols <- colnames(dplyr::select_if(confounder_summary, is.numeric))

DT::datatable(confounder_summary, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)

# Effect of sex, age and time together on causal effects in couples.

@@

- For each trait in the pipeline, we tested the causal effect within couples (i.e. the number of analyses corresponds to the number of traits (`r dim(traits_final)[1]`)). 
- The significant results after adjusting for multiple hypothesis testing are shown below. 
- `IVW_meta_beta` and `IVW_meta_pval` corresponds to the beta and p-value of the meta-analyzed MR across sexes, respectively (i.e. MR estimates were computed in each sex-seperately using sex-specific SNP-exposure and SNP-outcome results and then meta-analyzed).  
- After adjusting for multiple tests (p < `0.05/`r num_tests_by_PCs``), identified `r dim(AM_result_sig)[1]` significant assortative mating (same-trait) MR results (corresponding to the table below)..

@@

## Sex heterogeneity.

@@

## Binned results (exploring effect of age and time-together).

@@

slope_sig_age <- household_MR_binned_het_SNPmeta %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "age_even_bins") %>% 
  filter(bin_slope_pval < 0.05/num_tests_by_PCs_AM_sig)

slope_sig_tt <- household_MR_binned_het_SNPmeta %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "time_together_even_bins") %>% 
  filter(bin_slope_pval < 0.05/num_tests_by_PCs_AM_sig)

slope_wt_sig_age <- household_MR_binned_het_SNPmeta %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "age_even_bins") %>% 
  filter(bin_slope_pval_wt < 0.05/num_tests_by_PCs_AM_sig)

slope_wt_sig_tt <- household_MR_binned_het_SNPmeta %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "time_together_even_bins") %>% 
  filter(bin_slope_pval_wt < 0.05/num_tests_by_PCs_AM_sig)

@@

Given that we have no significant results, how should we report these analyses?

A few potential figures we could include for a given trait are shown below. 
```{r binned_figures, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = TRUE}

tar_load(exposure_info)
for(i in 1:length(exposure_info)){
  temp <- exposure_info[[i]]
  if(temp[which(temp$Value=="trait_ID"), "Info"] == "130_irnt") {break}
}

figs <- tar_read(household_MR_binned_AM_figs, branches = i)[[1]]

figs[[1]]
figs[[2]]
figs[[3]]
figs[[4]]
figs[[5]]
figs[[6]]
figs[[7]]
figs[[8]]
figs[[9]]

@@

# Heterogeneity stats

- In the single trait analysis, we also decided to investigate Cochran’s heterogeneity Q-stat to identify traits with high heterogeneity.
- This would indicate that associations between the index genotype and partner’s phenotype may not only act indirectly through causal relationship between the traits, but some direct effect is present. 
- Ideally, for the majority of the traits, the heterogeneity would be low, confirming that (despite what Tenesa suggests) there are no direct index genome to partner phenome effects.
- We tested heterogeneity statistics using the function [mr_heterogeneity](https://mrcieu.github.io/TwoSampleMR/reference/mr_heterogeneity.html) from the [TwoSampleMR](https://mrcieu.github.io/TwoSampleMR/index.html) package.

```{r, het_stots, include = FALSE}
sig_het <- household_MR_summary_AM %>% filter(Het_IVW_Qpval_round < 0.05/num_tests_by_PCs)
```

The number of traits with significant heterogeneity in the MR is: `r dim(sig_het)[1]`.

# Impact of confounders on couple correlation.

We tested the impact of confounding on genetic PCs, N/E coordinates and the following 4 traits:

```{r confounder_traits, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
confounder_trait_tibble <- as_tibble(confounder_traits)
confounder_trait_tibble_descript <- replace_Neale_ID(confounder_trait_tibble, traits_corr2_filled, "value", "outcome_description")
confounder_trait_tibble_descript %>% 
  kbl(col.names = NULL) %>%
  kable_styling()
```

## Ratio of confounding to correlation.

We tested the ratio of confounding to correlation for each confounder trait tested.

```{r corr_due_to_confounding_ratio, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

corr_impact_by_coords_ratio <- corr_impact_by_coords %>% 
  dplyr::select(Neale_pheno_ID, trait_couple_corr, trait_couple_corr_se, corr_due_to_confounding_all, corr_due_to_confounding_all_se) %>%
  unique() %>%
  mutate(trait_interest = "household_coords") %>%
  mutate(trait_ratio = corr_due_to_confounding_all/trait_couple_corr)

corr_impact_by_PCs_ratio <- corr_impact_by_PCs %>% 
  dplyr::select(Neale_pheno_ID, trait_couple_corr, trait_couple_corr_se, corr_due_to_confounding_all, corr_due_to_confounding_all_se) %>%
  unique() %>%
  mutate(trait_interest = "PCs") %>%
  mutate(trait_ratio = corr_due_to_confounding_all/trait_couple_corr)

corr_impact_by_traits_ratio <- corr_impact_by_traits %>% group_by(trait_interest) %>% 
  dplyr::select(trait_j, trait_interest, trait_j_couple_corr, trait_j_couple_corr_se, 
                corr_due_to_confounding, corr_due_to_confounding_se) %>% 
  mutate(trait_ratio = corr_due_to_confounding/trait_j_couple_corr)

corr_impact_by_traits_ratio <- replace_Neale_ID(corr_impact_by_traits_ratio, traits_corr2_filled, "trait_interest", "outcome_description") %>%
  dplyr::rename(trait_interest = outcome_description)

corr_impact_ratios <- rbind(corr_impact_by_PCs_ratio, corr_impact_by_coords_ratio, corr_impact_by_traits_ratio)
corr_impact_ratios_summary <- corr_impact_ratios %>%
  group_by(trait_interest) %>%
  dplyr::summarise(mean_ratio = mean(trait_ratio), median_ratio = median(trait_ratio), sd_ratio = sd(trait_ratio), number_of_traits = n())

corr_impact_ratios_summary %>%
  kbl() %>%
  kable_styling()

*A note on the number of traits:*

We tested the impact of `trait_interest` on couple correlation amongst all traits in our couple MR framework (number of traits = `r dim(household_MR_summary_AM)[1]`) with the exception of home location and birth place coordinates (2 each [N and E] * 2), we also removed the trait itself from the confounder analysis where applicable (in the case of traits `20016_irnt`, `738` and `845`). [Trait `189_irnt` was not included in our final couple MR analysis but could still be evaluated here because no MR-estimates were used.]

## Plots.

The figure below shows the 4 trait which showed the greatest impact on couple correlation (mean confounding ratio > 0.01).

@@

geography_plots <- list()

@@

  if(trait=="738"){trait_description <- "Average total household \nincome before tax"}
  if(trait=="189_irnt"){trait_description <- "Townsend deprivation index \nat recruitment"}

  plot_data <- corr_impact_by_traits %>% filter(trait_interest==trait) %>% filter(trait_interest!=trait_j)
  plot_data <- replace_Neale_ID(plot_data, traits_corr2_filled, "trait_j", "Phenotype")
  plot <-
    ggplot2::ggplot(
      data = plot_data,
      ggplot2::aes(x = trait_j_couple_corr,
                   y = corr_due_to_confounding, Phenotype = Phenotype)
    ) +
    theme(plot.title = element_text(hjust = 0.5)) +
    geom_point(alpha = 3 / 4) +

    ggplot2::geom_errorbar(
      ggplot2::aes(ymin = corr_due_to_confounding - corr_due_to_confounding_se, ymax = corr_due_to_confounding + corr_due_to_confounding_se),
      colour = "grey",
      width = 0
    ) +
    ggplot2::geom_errorbar(
      ggplot2::aes(xmin = trait_j_couple_corr - trait_j_couple_corr_se, xmax = trait_j_couple_corr + trait_j_couple_corr_se),
      colour = "grey",
      width = 0
    ) +

    theme_minimal() +
    ggplot2::labs(
      x = paste0("Couple trait correlation"),
      y = paste0("Couple correlation \ndue to confounding"),
      title = paste0(trait_description)
      #title = paste0("Impact of ", trait_description, "\n on couple correlation")
    ) +
    theme(plot.title = element_text(size=11, hjust = 0.5)) + theme(axis.title = element_text(size = 10))

@@

figures_grid <- plot_grid(plotlist = plots, nrow=2, ncol = 2, byrow = F, labels="AUTO")
figures_grid
```

# Comparison of paths from index to partner.

As a recap, the model is below.

```{r YiYp_adj_model, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/two_trait_model_yiyp_adj.png", error = FALSE)
```

We sought to compare the different paths from $X_i$ to $Y_p$ using the estimates computed as $\omega$, $\rho$ and $\gamma$.

```{r proxyMR_figure2, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
proxyMR_prod_comparison_fig_yiyp_adj
```

We considered the regression of the following estimates vs $\omega$ with an intercept passing through the origin (i.e. $estimate \sim \omega + 0$:

1. $\rho$
2. $\gamma$
3. $\rho$ + $\gamma$
4. $\rho_{resid} + \gamma$, where $\rho$  is residualized for $\gamma$ (i.e. $\rho_{resid} = resid(lm(\rho \sim \gamma))$)

The results are summarized below:

```{r regression_summary, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

data <- proxyMR_comparison_summary_yiyp_adj_SNPmeta
data$rho_resid <- resid(lm(rho_beta ~ gam_beta, data = data))
data$gam_rho_resid <- data$rho_resid + data$gam_beta

result <- numeric()
for(outcome in c("rho", "gam", "gam_rho", "gam_rho_resid")){

  col_to_rename <- paste0(outcome, "_beta")
  if(outcome=="gam_rho_resid") {col_to_rename <- outcome}
  data_model <- data %>% dplyr::rename(outcome = !!col_to_rename)
  model <- summary(lm(outcome ~ omega_beta + 0, data = data_model)) 
  r2 <- model %>% glance %>% pull(r.squared)
  model_summary <- model %>% tidy() %>% 
    mutate(term = "omega") %>% mutate(outcome = outcome) %>% mutate(r.squared = r2)

  result <- rbind(result, model_summary)

}

result %>% dplyr::select(outcome, everything()) %>%
  kbl() %>%
  kable_styling()

@@

```{r remove_opp_sign, include = FALSE}

opp_sign_rows <- numeric()
for(col1 in c("gam_beta", "rho_beta", "gam_rho_beta", "gam_rho_resid", "omega_beta")){

  for(col2 in c("gam_beta", "rho_beta", "gam_rho_beta", "gam_rho_resid", "omega_beta")){

      temp <- which(sign(data[[col1]])!=sign(data[[col2]]))
      opp_sign_rows <- c(opp_sign_rows, temp)
  }

}

opp_sign_rows <- unique(opp_sign_rows)

To further explore these estimates, performed paired t-tests to find the significant differences between pairs of estimates. We first removed all instances where the sign did not match between any two pairings. In total, of the `r dim(data)[1]` trait pairs under consideration, `r length(opp_sign_rows)` were removed.

```{r boxplot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}

df <- data[-opp_sign_rows,] %>% dplyr::select(exposure_ID, outcome_ID, gam_beta, rho_beta, gam_rho_beta, gam_rho_resid, omega_beta)

data_boxplot  <- df %>%
  pivot_longer(cols = !c("exposure_ID", "outcome_ID"), names_to = "data", values_to = "beta") %>%
  mutate_at("data", as.factor) %>%
  mutate(
    data = fct_recode(as.factor(data),
                      Gamma = "gam_beta",
                      Rho = "rho_beta",
                      Gamma_Rho = "gam_rho_beta",
                      Gamma_Rho_resid = "gam_rho_resid",
                      Omega = "omega_beta")
  )

pairwise.t.test(data_boxplot$beta, data_boxplot$data, p.adjust = "none")

cmpr <- list(c("Omega", "Gamma_Rho"), c("Gamma","Gamma_Rho"), c("Gamma_Rho","Gamma_Rho_resid"), c("Rho", "Gamma_Rho"))

position <- data_boxplot %>% group_by(data) %>% dplyr::summarise(median=median(beta)) %>% 
  arrange(-desc(median)) %>% pull(data)

data_boxplot$data <- factor(data_boxplot$data,levels=position)

boxplot <- data_boxplot %>%
  ggplot( aes(x=beta, y=data)) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
  theme_ipsum() +
  theme(
    legend.position="none",
    plot.title = element_text(size=11)
  ) +
  labs(x ="Estimate", y = "") +
  scale_y_discrete(labels=c("Gamma_Rho" = "\u03B3 + \u03C1", "Omega" = "\u03C9",
                            "Gamma" = "\u03B3", "Gamma_Rho_resid" = expression( paste(rho ["resid"], " + ", gamma)), "Rho" = "\u03C1")) +
  stat_compare_means(comparisons = cmpr, tip.length=0.01,
                     label = "p.signif",
                     symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1),
                                        symbols = c("****", "***", "**", "*", "ns")))

boxplot
